---
title: "R Notebook"
output: html_notebook
---

Overview:
This code maps and extracts HSI and RGB from the NEON site CPER

CPER: "full" site, 2024 (tiles where there are also NEON TOS sampling plots)
```{r}
#input with paths is all produced tiffs of tiles:
input_with_paths <- read.csv('CPER_2024_tiles.csv')
hsi_list <- unique(input_with_paths$hsi_saved_path)
rgb_list <- unique(input_with_paths$rgb_filepath)
spbands_list <- unique(input_with_paths$spbands_saved_path)

#mapping rgb
input <- rgb_list %>% na.omit() #for every rgb tile file input
#tiles_only = just map and save the tiles in site_tilelist; 
# if tiles_only == FALSE, we also clip out PLOTS
tiles_only <- TRUE
site_tilelist <- c()

foreach(t = 1:length(input),
        .combine = rbind) %do% {
          
          this_tile <- terra::rast(input[[t]])
          tile_extent <- ext(this_tile)
            
          if (tiles_only == TRUE) {
            #saves a list of them for full site mapping
            site_tilelist[[t]] <- this_tile
          }
          
          #default map:
          tilemap <- terra::plotRGB(this_tile) #r = 1, g = 2, b = 3 is default
          
          #dramatic map:
          tilemap <- terra::plotRGB(this_tile, r=1, g=2, b=3, stretch="lin")
          
          
          if (tiles_only == FALSE) {
            
            plots_projected <- terra::project(site_subplots_shp, this_tile) #can also use David's data here
            plots_in_tile <- plots_projected[tile_extent]
            
            if (nrow(plots_in_tile) > 0) { #if there are plots in a tile, create lookup table
              #create a translation between extraction ID and plotID
              plot_lookup <- data.frame(
                ID = 1:nrow(plots_in_tile), #we will make a dataframe with plot ID (just a number)
                subplotID = plots_in_tile$subplotID, #the subplot and plot ID (from the dataframe)
                plotID = plots_in_tile$plotID #so that we can reference which plot is which
              )
              
              hello <- foreach(ID = nrow(plots_in_tile),
                               .combine = rbind) %do% {
                                 
                                 this_plot <- plots_in_tile[ID,] #for each plot, numbered 1 to nrow
                                 
                                 this_plot_df <- as.data.frame(this_plot)
                                 subplotID = this_plot_df$subplotID
                                 
                                 buff_size = 0.5*sqrt(this_plot_df$subpltSize) +2
                                 buff_extent <- terra::ext(terra::buffer(this_plot, 
                                                                         width = buff_size))
                                 cropped_plot <- terra::crop(this_tile, buff_extent) #crop the map to just the buffered point
                                 
                                 terra::plotRGB(cropped_plot)
                                 terra::plotRGB(cropped_plot, r = 1, g = 2, b = 3, stretch = 'lin') #plot the cropped plot
                                 
                                 this_plot_df
                               }
            } #plots in tile
            hello
          } #tiles only = FALSE
          
        } #each file

#create full site rgb image:
full_site_rgb10 <- terra::merge(terra::sprc(site_tilelist))
full_site_dullrgb <- terra::plotRGB(full_site_rgb10)
full_site_stretchrgb <- terra::plotRGB(full_site_rgb10, stretch = 'lin')

```



```{r}
#need to create veg index with spbands_list:
input <- spbands_list
VI_list <- c() #initialize empty list
#using spbands_list, make vegetation index rasters and an RGB composite: 
foreach(t = 1:length(input)) %do% {
  this_tile <- terra::rast(input[t])
  tile_extent <- ext(this_tile)
  
    #make rasters
    PRI_rast <- (this_tile[['NIR']] - this_tile[['green']]) / (this_tile[['NIR']] + this_tile[['green']])
    NDVI_rast <- (this_tile[['NIR']] - this_tile[['red']]) / (this_tile[['NIR']] + this_tile[['red']] + 0.0001)
    
    NIRv_rast <- this_tile[['NIR']] * NDVI_rast
    CCI_rast <- (this_tile[['PRI']] - this_tile[['red']]) / (this_tile[['PRI']] + this_tile[['red']])
    
    indices <- c(CCI_rast, NIRv_rast, PRI_rast)
    names(indices) <- c("CCI", "NIRv", "PRI")

    VI_list[[t]] <- indices  
}

  NIRv_col = colorRampPalette(c("white", "darkgreen"))(20)
  PRI_col = colorRampPalette(c("white", "blue"))(20)
  CCI_col = colorRampPalette(c("white", "red"))(20)

input <- VI_list
foreach(t = 1:length(input)) %do% {
  this_tile <- VI_list[[t]]
  terra::plot(this_tile$CCI, col = CCI_col, main = "CCI")
  terra::plot(this_tile$NIRv, col = NIRv_col, main = "NIRv")
  terra::plot(this_tile$PRI, col = PRI_col, main = 'PRI')

  VI_rast <- terra::plotRGB(this_tile, r=1, g=2, b=3, stretch="lin")
}

#create fullsite vi map:
full_site_vis <- terra::merge(terra::sprc(VI_list))
terra::plotRGB(full_site_vis, r = 1, g = 2, b = 3, stretch = 'lin')
```



```{r}
rgb_map_sampling_plot('CPER')
```
```{r}
VI_sampling_plot( 'CPER')

```

```{r}

cov_specific_char <- characterize_veg(general = FALSE, species_only = FALSE)

#specific cover (mat_spec)
  #PC1: + standing dead herbaceous; - soil; - Salsola, Plantago, Amaranthera?
  #PC2: only positive: litter, Sisymbrium altissium; less so vulpia octoflora, Elymus elymoides, Lappula occidentalis, Lepidium densiflorum.
  #PC3: most negative: lappula occidentalis, Sisymbrium altissium. lepidium densiflorum
  
  # + PC1: hella standing dead herbaceous, little soil, Salsola, Plantago, Amaranthera.
  # - PC1: hella soil and little standing dead herbaceous. Maybe also hella Salsola, Plantago, and Amaranthera?
  
  # + PC2: lots of litter and sisymbrium
  # - PC2: little litter and sisymbrium
  
  # + PC3: little lap occ, sisymbrium altissium, and lepidium densiflorum.
  # - PC3: lots of lap occ, sisymbrium altissium, and lepidium densiflorum.
  
cov_specific_plotsums <- cov_specific_char %>%
  mutate(plot = strtrim(subplot, 8)) %>%
  group_by(plot, PC) %>%
  summarise(plot_pc_loc = mean(subplot_loc))

cov_specific_plotsums %>%
    pivot_wider(names_from = PC, values_from = plot_pc_loc) %>%
  ggtern(aes(x = PC1, y = PC2, z = PC3)) +
  geom_point(aes(colour = plot), size = 3, alpha = 0.5) +
  #xlim() +
  #ylim() +
  #zlim() +
  theme_classic()

print()
```
  mutate(relationship = ifelse(plot_pc_loc < 0, 'negative', 'positive'))
  
cov_specific_plotsums %>%
  mutate(plot_pc_loc = abs(plot_pc_loc)) %>%
  pivot_wider(names_from = PC, values_from = c(relationship, plot_pc_loc)) %>%
  print()
                
  ggtern(aes(x = PC1, y = PC2, z = PC3)) +
  geom_point(aes(colour = plot), size = 3, alpha = 0.5) +
  theme_classic()

posi_rels <- cov_specific_plotsums %>%
  filter(relationship == 'positive') %>%
neg_rels
posi_rels
```


```{r}
cov_gen_char <- characterize_veg(general = TRUE, species_only)

```
#general cover (mat_gen)
  # shows us that PC1 is strongly affected by... 1) soil (positively), 2) standing dead (negatively), plants (negatively), and 3) rock (positively)
  # PC2 is strongly affected by litter, wood (positively) and negatively with lichen and moss
  # PC3 affected by fungi (negatively), plants (positively)
  
  # + PC1 = hella soil, some rock; little standing dead or plants
  # - PC1 = lots of standing dead or plants
  
  # + PC2 = hella litter and wood, little lichen and moss
  # - PC2 = hella lichen and moss, little litter and wood
  
  # + PC3 = plants
  # - PC3 = fungi
  
```{r} 
# capturing vegetation “vibe” using % cover
cov_species_char <- characterize_veg(general = FALSE, species_only = TRUE)
```


#show clustering by species groups hopefully

#consider how this might translate to a map, spatial / grid patterns added to what we already see

  # Summarize: 
  # Whole plot, subplot size-specific analyses of % cover
  # relative frequency for plot and larger subplots
  
  
  # among (matrix rows) in terms of cover and composition, 
  #which is then used to characterize each organizational level: 
  # [a diagram of subplot, plot, site, domain, all NEON sites]
  #and what we're dealing w here...
  

# vegetation vibe: PA, % cover

#whole plot summary of subplots

# subplot specific pairings

# capturing vegetation "vibe" using PA

```




```{r}

extracted_100m <- extract_spectra(hsi_list, 'CPER', 100)

  terra::plot(plots_in_tile, add = TRUE, lwd = 3, col = 'white')
  
#layout for 1, 1, 2, 3; 1, 1, 4, 5
  layout_matrix <- matrix(c(1, 1, 2, 3,
                          1, 1, 4, 5), 
                        nrow = 2, byrow = TRUE)

# 2. Apply the layout
layout(layout_matrix)


terra::plotRGB(VI_rast, r = 1, g = 2, b = 3, stretch = 'lin')
terra::plot(plots_in_tile, add = TRUE, lwd = 3, col = 'white')

```
These are DATA-RICH maps of tiles, where the R channel is CCI, G is NIRv, and B is PRI
That means if something is very GREEN it has high NIRv
If it's very red, high CCI, very blue, high PRI; 
CCI = long-term stress indicator OR senescence
PRI = additional stress indicator but can be shorter or longer term.

The white squares are where sampling plot locations are marked in the tile.
The smallest sized squares are veg plots